---
title: "Predicting Hotel Cancellations"
author: "Art Tay"
output: pdf_document
header-includes:
    - \usepackage{caption}
format:
  pdf:
     documentclass: article
     papersize: letter

execute:
  warning: false
  message: false
---
\captionsetup[table]{labelfont = bf}
\captionsetup[figure]{labelfont = bf}
```{r setup, include = FALSE}
##Setup code
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
knitr::opts_knit$set(eval.after = "fig.cap")

# Libraries
library(magrittr)
library(tidyverse)
library(tidymodels)
library(naniar)
library(cutpointr)
library(kableExtra)
library(vip)
library(MASS)
library(ggpubr)
library(DescTools)
tidymodels_prefer()
```

```{r Parallel Setup, include = F}
library(doParallel)
library(tictoc)
cl <- makePSOCKcluster(24)
registerDoParallel(cl)
```

# Abstract
Hotels are extremely interested in predicting room cancellations as
utilized capacity is a key driver of revenue. Data originally complied
by *Antonio et. Al, 2019*, on over 100,000 hotel bookings was appropriated
to build several logistic regression models. The final model had a predictive
accuracy on a held-out test set of 80.5%. In conjunction with exploratory
data analysis inferences beyond prediction were drawn. First, although the
data was complied from two different types of hotels, and single predictive model
was more powerful than a split model. This applies that the same factors that
predict cancellations at Resort hotel can be used for City hotel and possibly other
types of hotels. Second, all the feature present in the dataset were found to
be useful predictors. On a forward basis, hotels should consider collecting
and maintain a similar database of customer data. Further investigation is
required to identify exigent circumstances effect Portugal bookings during the
fall of 2015 as well as the rational behind canceled bookings where a non-refundable
deposit was made.

# Introduction
One of the primary concerns of a hotel is guests that end up canceling reservations.
Cancelled reservations can dramatically impact revenue by reducing the overall
utilization rate of rooms. *Antonio et Al., 2019*, collected bookings data from a
resort hotel and a city hotel from a two year period. Over 100,000 bookings
are represented in the dataset. The dataset contains categorical information
about each booking such as the agent it was booked through, the date of arrival,
and whether or not it was ultimately cancelled. The dataset also contained
numeric information such as the number of previous cancellation a guest has
made, and the number of guests on the bookings. Given the growing need to understand
hotel cancellations, the goal of this analysis was to build a model that could
predicted which books are most likely to be cancelled.

\pagebreak

# EDA
```{r class imbalance tabulation, echo = F, fig.cap = caption}
#| out.width = '60%',
#| fig.align = 'center'
load(file = "Figures/marimekko_plot.rds")
caption <- "Number of bookings grouped by hotel type and cancellation status.
            The height on each box represents the percentage of bookings
            by cancellation status, and the width the percentage of bookings
            made in each type of hotel."
marimekko_plot
```

70% of the data was extract to conduct the exploratory data analysis as to not
bias the final error rates. Based on this subset, the number of cancelled reservations
is quite balanced with the number of uncancelled reservation. The figure above
indicates that the data is more imbalance for Resort Hotels, but the ratio in
neither class exceeds 4:1.

```{r Figure 1, echo = F, fig.cap = caption}
#| out.width = '60%',
#| fig.align = 'center'
load(file = "Figures/plots_tables.rds")
caption <- "Percentage differences in means of numeric predictors
  for Cancelled v. Uncancelled reservations. Percentages were calculated
  relative to Cancelled reservations. Positive value indicate that the
  mean value of the Cancelled group is greater than the mean of the
  Uncancelled group."
plot_1
```

The plot above clear demonstrates the that guests who have previously canceled
a bookings are far more likely to cancel another. Guest that cancelled reservations
had nearly 1,000% more previous cancellation than guests that checked-in. Less
significant indicators of cancellation included a longer lead time and a greater
number of days on the waiting list.

On the contrary, guests that make special
requests or require additional parking spaces are less likely to cancel.
This may be caused by prepaid fee associated with these types of requests.
Interesting the duration and type of day that booking covered did not
differ between cancelled and uncanceled reservations. Long stay over
a weekend seem just as likely to be cancelled as a short business trip during
the week.

```{r factor plot, echo = F, fig.cap = caption}
#| out.width = '60%',
#| fig.align = 'center'
load("Figures/factor_plot.rds")
caption <- "Percentage of different categories
of bookings colored by cancellation status."
factor_plot
```

Although many different types of bookings are represented in the data,
there is little separation in proportion that get cancelled. Bookings in made
Portugal are cancelled at a higher rate than any other country. The majority
of these cancellation happened during the fall of 2015, which may indicate an
event effected the country resulting cancellations; however, there were other
time period with high rates of cancellations from Portugal. Strangely, the
data seem to indicate that reservations that were made with a non-refundable
deposit are more likely to be canceled. Smaller effects that can be seen
on the above plot are that Groups are more likely to cancel, while, direct
bookings are less likely to be cancelled.

```{r Figure 2, echo = F, fig.cap = caption}
#| out.width = '60%',
#| fig.align = 'center'
caption <- "Percentage differences in means of numeric predictors
  for City v. Resort hotels. Percentages were calculated relative to
  the City group. Positive values indicate that the mean value of the City
  group is greater than the mean of the Resort group."
plot_2
```

\pagebreak

The plot above indicates that there is a much higher demand for City Hotels than
there is for Resort Hotels as client spend almost 500% more time on the waiting
list for City Hotels. This is probably a result of seasonality. City Hotels are
used year around for business trips, whereas Resort Hotels are booking more
infrequently. On the other hand, resort bookings tend to require more special
accommodations.

```{r factor plot hotel, echo = F, fig.cap = caption}
#| out.width = '60%',
#| fig.align = 'center'
load("Figures/factor_plot_hotel.rds")
caption <- "Percentage of different categories of bookings
colored by type of hotel"
factor_plot_hotel
```
Similar to the plot coded by cancellation status, Figure 5 indicates that
although a variety of bookings types are represented in the data, there
is very little separation on the basis of hotel.

# Modeling Methodology
The data was initially split into a training data set and a testing data set
(70% and 30% of the raw data respectively). Although some factors had a NULL
level, it was generally meaningful as opposed to missing. For example, the
majority of the agent data was NULL, but this was taken to mean that the
customer simply booked a hotel without using an agent.

The `reservation_status`
as well as `reservation_status_date` were both dropped from the feature space.
The `reservation_status` variable was perfectly correlated with the outcome
variable, the cancellation status, and therefore unlikely to be known prior
to the actual cancellation. Although `reservation_status_date` is not perfectly
correlated to the outcome, it is impossible to separate the intermediate
status updates from the final outcome. The inclusion of this predictor in the
model would most likely result in a high optimistic error rate.

Additional,
all factors were converted into $n-1$ dummy variables. Then, near-zero-variance
predictor as well as highly correlated variables were dropped. Finally, all
originally numeric variables were centered and scaled to promote computational
stability.

This data was then used to fit a standard logistic regression model,
a penalized logistic regression model, and finally a split model based on hotel.
The final model was selected on the basis of test set accuracy. Accuracy was
calculated using the probability cut-point that maximized the kappa statistic
on the training set for each given model. No models had Cook's Distance value
greater than 1, therefore no observations were removed from the training sets
of any of the models. Although regularization was applied, the optimal penalty
was not significantly different from 0, and thus was unreported.

\pagebreak

# Results
```{r, echo = F}
load("Figures/table_5.rds")
table_5_latex
```
The overall accuracy of the standard logistic regression model (combined model)
was $0.815$. Model metrics were revaluate by hotel to make a direct comparison
to the split model. Although, the metrics are close, the combined model
dominates the split model in accuracy across both groups. This is mostly likely
to the higher data utilization of the combined model. This might also be the
result of the two hotels having a very similar distribution of features.

```{r Figure 3, echo = F, fig.cap = caption}
#| out.width = '70%',
#| fig.align = 'center'
load(file = "Figures/plot_4.rds")
caption <- "Ten Most Important Variable for the Final Model"
plot_4
```

Based on the variable importance plot, the most important categorical variable
was country, and the most important numeric variable was the number of special
requests. Based on the sign of the associated coefficients  bookings from
Portugal were more likely to be cancel, while the more special requests a
customer made the less likely they were to cancel (for a full list of
model coefficients refer the the *Appendix*). Interesting various agent
variables were flag as significant to the model despite the significant dispersion
of agents with most having a low level of separation between the canceled and
uncanceled bookings.

# Conclusion
The overall goal of this analysis was to produce a predictive model that could
estimate the probability on a client cancelling a bookings with a high degree
of accuracy. The ultimate model chosen as a standard logistic regression model.
The final model had a test accuracy of 80.5%. During the exploratory data analysis
the implication of several variables on the ultimate outcome were hypothesized.
Many on these hypothesizes were consistent with the model coefficients. For
example the model predicts a greater cancellation rate for customers with a
high number of previous cancellations. There are two issues that may require
further investigation. Firstly, there were a high number of cancellations
from Portugal made during the fall of 2015. It is possible that this trend
will continue into the future, but exigent circumstances should be identified
to increase the applicability of the model on future data. Second, the data seems
to indicate that bookings made with a non-refundable deposit are more likely to
be canceled. This seems logical backwards, as it is usually the case that
customers are less likely to forego a partial prepaid room.

\pagebreak

# Appendix - Additional Plots
```{r, echo = F}
load("Figures/model_1_coef.rds")
model_1_coef
```

# Appendix - Code

## Project Description

* Perform an EDA – Brief EDA on that hotel booking data (city room vs. resort)
* Build a logistic regression model if one would cancel a reservation or not
* Check class imbalance
* Provide an evaluation Metric
* Show the feature needed any hot encoding

## EDA
* Slight class imbalance. Roughly 2:1 Uncancelled:Cancelled. Unlike to cause
modeling problems.
* Some variable difference identified between outcome classes.
* Slight differences between variables between hotel groups. Most likely
insufficient to justify separate models.

```{r load Data}
data_full <- read_csv(file = "hotel_bookings.csv")
```

```{r check data}
#| eval: false
str(data_full)
glimpse(data_full)
```

### Data Cleaning
```{r Train Test Split}
# Train-Test Split (70/30)
set.seed(123)
data_split <- initial_split(data_full, prop = 0.7)

train <- training(data_split)
test <- testing(data_split)
```

```{r Initialized Recipe}
# start a recipe
cleaning_recipe <- train %>%
                   recipe(is_canceled ~ .)
```

```{r Data Type Coding}
# Change all character variables to factors.
cleaning_recipe %<>%
  step_mutate(across(where(is.character), as.factor))

# Change outcome variable to be a factor.
cleaning_recipe %<>%
    step_mutate(is_canceled = as.factor(is_canceled), skip = T)

# Change month string to corresponding month number
cleaning_recipe %<>%
    step_mutate(arrival_date_month = match(arrival_date_month, month.name))

# Change arrival (date) variables to factors.
cleaning_recipe %<>%
    step_mutate_at(starts_with("arrival"), fn = as.factor)

# Change is_repeated_guest to be a factor variable.
cleaning_recipe %<>%
    step_mutate(is_repeated_guest = as.factor(is_repeated_guest))
```

```{r intermediate bake}
#| include = F
cleaning_recipe_prepped <- cleaning_recipe %>% prep(retain = T, verbose = T)
train_clean <- bake(cleaning_recipe_prepped, new_data = NULL)
str(train_clean)
```

```{r missing values, include = F}
miss_count <- sapply(train_clean, FUN = function(x) {sum(is.na(x))})
```

### Tables
```{r EDA tables 1 and 2}
mean_not <- train_clean %>% filter(is_canceled == 0) %>%
                           summarise(across(where(is.numeric), mean)) %>%
                           round(digits = 2)

mean_canceled <- train_clean %>% filter(is_canceled == 1) %>%
                           summarise(across(where(is.numeric), mean)) %>%
                           round(digits = 2)

table_1 <- as.data.frame(t(rbind(mean_canceled, mean_not)))
colnames(table_1) <- c("Uncancelled", "Cancelled")
table_1 %<>% mutate("%_Diff" = round(Uncancelled / Cancelled - 1, 2))

table_2 <- train_clean %>%
           group_by(is_canceled) %>%
           summarize(across(where(is.factor), Mode)) %>%
           t() %>% as.data.frame()

colnames(table_2) <- c("Uncancelled", "Cancelled")

table_2 %<>% filter(Uncancelled != Cancelled)
```

```{r EDA tables 3 and 4}
mean_city <- train_clean %>% filter(hotel == "City Hotel") %>%
                           summarise(across(where(is.numeric), mean)) %>%
                           round(digits = 2)

mean_resort <- train_clean %>% filter(hotel == "Resort Hotel") %>%
                           summarise(across(where(is.numeric), mean)) %>%
                           round(digits = 2)

table_3 <- as.data.frame(t(rbind(mean_city, mean_resort)))
colnames(table_3) <- c("City", "Resort")
table_3 %<>% mutate("%_Diff" = round(City / Resort - 1, 2))

table_4 <- train_clean %>% group_by(hotel) %>%
                           summarize(across(where(is.factor), Mode)) %>%
                           t() %>% as.data.frame()

colnames(table_4) <- c("City", "Resort")

table_4 %<>% filter(City != Resort)
```

```{r Marimekko Plot}
marimekko_data <- train_clean %>%
                  # 2 by 2 count table
                  count(hotel, is_canceled) %>%
                  group_by(hotel) %>%
                  # percentage of cancelled in each hotel group
                  mutate(Percent_row = n * 100 / sum(n), n = NULL) %>%
                  pivot_wider(names_from = is_canceled,
                              values_from = Percent_row) %>%
                  ungroup() %>%
                  # Percent each hotel of the entire dataset
                  mutate(Percent_col = c(sum(`0`), sum(`1`))) %>%
                  # Define xmax and xmin for rectangle plots
                  mutate(Percent_col = Percent_col * 100 / sum(Percent_col)) %>%
                  mutate(xmax = cumsum(Percent_col),
                         xmin = xmax - Percent_col) %>%
                  pivot_longer(col = c(`0`, `1`)) %>%
                  # Define ymax and ymin for rectangle plots
                  group_by(hotel) %>%
                  mutate(ymax = cumsum(value), ymin = ymax - value) %>%
                  # Define text positions for labels
                  mutate(xtext = xmin + (xmax - xmin) / 2,
                         ytext = ymin + (ymax - ymin) / 2) %>%
                  mutate(name = ifelse(name == 0, "Uncancelled", "Cancelled")) %>%
                  ungroup()

#marimekko_data

# 2 by 2 counts for labels
counts <- train_clean %>% count(hotel, is_canceled) %>% select(n)

marimekko_plot <- marimekko_data %>% mutate(counts = counts$n) %>%
ggplot(
  aes(ymin = ymin, ymax = ymax, xmin = xmin, xmax = xmax, fill = name)
) + geom_rect(col = I("grey")) +
    geom_text(aes(x = xtext, y = ytext,
              label = scales::comma(counts))) +
    geom_text(aes(x = xtext, y = 103, label = hotel), size = 4) +
    labs(x = "Percentage of Bookings by Hotel",
         y = "Percentage of Bookings by Cancellation Status",
         fill = "") +
    theme_bw() +
    theme(legend.position = "bottom")

marimekko_plot

save(file = "Figures/marimekko_plot.rds", marimekko_plot)
```

```{r Factor Plot by Cancelled}
train_factor_plots <- train_clean %>%
                      select(is_canceled, is_repeated_guest, customer_type,
                             deposit_type,
                             distribution_channel,
                             market_segment,
                             country)

counts_by_level <- function(colname){
  train_factor_plots %>% group_by(is_canceled) %>%
                  count(!!as.name(colname)) %>%
                  mutate(Percent = n * 100 / sum(n)) %>%
                  mutate(Factor = colname) %>%
                  rename(Level = !!as.name(colname)) %>%
                  arrange(desc(Percent)) %>%
                  slice(1:4)
}

names <- colnames(train_factor_plots %>% select(-is_canceled))

train_factor_plot_data <- do.call("rbind", lapply(names, FUN = counts_by_level))

factor_plot <- ggplot(data = train_factor_plot_data,
  aes(x = reorder(Level, -Percent), y = Percent, fill = is_canceled)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(vars(Factor), scales = "free_x") +
  theme_bw() +
  labs(x = "") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_fill_discrete(labels = c("Uncanceled", "Canceled")) +
  theme(legend.title = element_blank(), legend.position = "bottom")

factor_plot
```

```{r, check PRT}
train_clean %>% select(arrival_date_year,
  arrival_date_month, country, is_canceled) %>%
  filter(country == "PRT" & is_canceled == 1) %>%
  count(arrival_date_year, arrival_date_month) %>%
  arrange(desc(n)) %>% print(n = 26)
```

```{r Factor Plot by hotel}
train_factor_plots <- train_clean %>%
                      select(is_repeated_guest, customer_type,
                             deposit_type, hotel,
                             distribution_channel,
                             market_segment,
                             country)

counts_by_level_hotel <- function(colname){
  train_factor_plots %>% group_by(hotel) %>%
                  count(!!as.name(colname)) %>%
                  mutate(Percent = n * 100 / sum(n)) %>%
                  mutate(Factor = colname) %>%
                  rename(Level = !!as.name(colname)) %>%
                  arrange(desc(Percent)) %>%
                  slice(1:4)
}

names <- colnames(train_factor_plots %>% select(-hotel))

train_factor_plot_hotel <- do.call("rbind",
  lapply(names, FUN = counts_by_level_hotel))

factor_plot_hotel <- ggplot(data = train_factor_plot_hotel,
  aes(x = reorder(Level, -Percent), y = Percent, fill = hotel)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(vars(Factor), scales = "free_x") +
  theme_bw() +
  labs(x = "") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  theme(legend.title = element_blank(), legend.position = "bottom")

factor_plot_hotel
```

```{r}
save(file = "Figures/factor_plot.rds", factor_plot)
save(file = "Figures/factor_plot_hotel.rds", factor_plot_hotel)
```

```{r Check reservation_status, eval = F}
train_clean %>% group_by(is_canceled, reservation_status) %>%
                summarise(n = n())
train_test_perfect <- train_clean %>% select(is_canceled, reservation_status)

library(fastDummies)
train_test_perfect %<>% dummy_cols(remove_first_dummy = T,
                                   remove_selected_columns = T)

test_glm <- glm(is_canceled_1~ ., family = binomial,
                    data = train_test_perfect)

train_test_perfect$pred <- predict(test_glm, type = "response")
train_test_perfect %<>% mutate(pred = ifelse(pred >= 0.5, 1, 0))

train_test_perfect %>% ggplot(
  aes(x = is_canceled_1, y = pred, col = "blue")
) + geom_point()
```

### Plots
```{r Plot 1}
plot_1 <- table_1 %>%
   mutate("variables" = rownames(.), "diff" = `%_Diff` * 100) %>%
   filter(variables != "children") %>%
  ggplot(
    aes(x = diff, y = variables, fill = variables)
  ) + geom_bar(stat = 'identity') +
      xlim(-100, 960) +
      labs(y = "", x = "Percent Difference in Means") +
      theme_bw() + theme(legend.position = "none",
                         axis.text.x = element_text(hjust = 1))
plot_1
```

```{r plot 2}
plot_2 <- table_3 %>%
   mutate("variables" = rownames(.), "diff" = `%_Diff` * 100) %>%
   filter(variables != "children") %>%
  ggplot(
    aes(x = diff, y = variables, fill = variables)
  ) + geom_bar(stat = 'identity') +
      labs(y = "", x = "Percent Difference in Means") +
      theme_bw() + theme(legend.position = "none")

plot_2
```

```{r table 2 formatting}
caption <- "Factor modes of Cancelled v. Uncancelled reservations"
table_2_latex <- table_2 %>%
  kbl(format = 'latex', booktab = T, longtable = T, caption = caption,
    linesep = "", escape = T) %>%
    kable_styling(full_width = F)
```

```{r table 4 formatting}
caption <- "Factor modes of City v. Resort hotels"
table_4_latex <- table_4 %>%
  kbl(format = 'latex', booktab = T, longtable = T, caption = caption,
    linesep = "", escape = T) %>%
    kable_styling(full_width = F)
```

```{r output for report use, include = F}
save(file = "Figures/plots_tables.rds",
  list = c("plot_1", "plot_2", "table_2_latex", "table_4_latex"))
```

## Model Fitting
```{r Model Specific Preprocessing}
# Guard against multicollinearity and overfitting.
# Not concerned about normality of skewness for logistic regression.
cleaning_recipe %<>% step_rm(reservation_status, reservation_status_date) %>%
                     step_dummy(all_nominal_predictors()) %>%
                     step_nzv(all_predictors()) %>%
                     step_corr(all_numeric_predictors()) %>%
                     step_normalize(all_numeric_predictors())
```

### Model 1
```{r Define Model Workflow}
log_Reg_mod <- logistic_reg() %>% set_engine("glm")

log_Reg_wflow <- workflow() %>%
                 add_model(log_Reg_mod) %>%
                 add_recipe(cleaning_recipe)

log_Reg_wflow
```

```{r Model 1 Fit}
# Saved and reloaded as rds to save runtime.
#log_Reg_fit <- log_Reg_wflow %>%
               #fit(data = train)
```

```{r save and load fit, include = F}
#save(file = "Figures/log_Reg_fit.rds", log_Reg_fit)
load("Figures/log_Reg_fit.rds")
```

```{r Model Summary}
log_Reg_glm <- log_Reg_fit %>% extract_fit_parsnip()

#summary(log_Reg_glm)

#par(mfrow = c(2, 2))
#plot(log_Reg_glm$fit)#, which = c(1,2))
```

```{r Influential Plot Analysis, eval = F}
Id <- 1:length(train$hotel)
Leverage <- hatvalues(log_Reg_glm$fit)
StudRes <- studres(log_Reg_glm$fit)
CookD <- cooks.distance(log_Reg_glm$fit)

inful_data <- cbind(Id, Leverage, StudRes, CookD)
inful_data <- as.data.frame(inful_data)

##Plots
##Leverage
lev <- ggplot(data = inful_data, aes(x = Id, y = Leverage)) + geom_point() +
            geom_hline(yintercept = 2 * length(log_Reg_glm$fit$coefficients) /
                length(inful_data$Id), col = "red") +
            labs(x = "Index") +
            theme_bw()

##Studentized Residuals
studres <- ggplot(data = inful_data, aes(x = Id, y = StudRes)) + geom_point() +
                geom_hline(yintercept = 2, col = "red") +
                geom_hline(yintercept = -2, col = "red") +
                labs(y = "Studentized Residuals", x = "Index") +
                theme_bw()

##Cooks distance
cooks <- ggplot(data = inful_data, aes(x = Id, y = CookD)) + geom_point() +
             geom_hline(yintercept = 1, col = "red") +
             labs(y = "Cook's Distance", x = "Index") +
             theme_bw()

inful <- ggarrange(lev, studres, cooks, ncol = 3, nrow = 1)

inful <- annotate_figure(inful, top = text_grob("Influential Point Analysis"))

inful

save(inful, file = "Figures/inful.rds")
```

### Model 2 (Remove Outliers)
```{r Check Cooks Distances, eval = F}
train_2 <- train %>%
    mutate(m1cooksD = cooks.distance(log_Reg_glm$fit))
dim(train_2)
train_2 %<>% filter(m1cooksD < 1)
```

### Model 3 (Split Model)
```{r Data Subset for Split Model}
# Define data splits
train_city <- train %>% filter(hotel == "City Hotel")
train_resort <- train %>% filter(hotel == "Resort Hotel")
test_city <- test %>% filter(hotel == "City Hotel")
test_resort <- test %>% filter(hotel == "Resort Hotel")
```

```{r Update workflow for model 3}
recipe_3 <- cleaning_recipe
rm_hotel <- train %>% recipe(formula = is_canceled ~ . ) %>%
                      step_rm(hotel)
recipe_3$steps <- append(recipe_3$steps, list(rm_hotel$steps[[1]]), after = 6)

model_3_wflow <- workflow() %>%
                  add_model(log_Reg_mod) %>%
                  add_recipe(recipe_3)
```

```{r Model 3 Fits, cache = T}
model_3_fit_city <- model_3_wflow %>% fit(data = train_city)
model_3_fit_resort <- model_3_wflow %>% fit(data = train_resort)
```

### Model 4 (Penalized)
```{r Model 4 Tunning}
# penalty = lambda
# mixture = alpha
# Define elastic net model.
pen_log_mod <- logistic_reg(penalty = tune(), mixture = tune()) %>%
               set_engine("glmnet")

# Define resampling method for hyperparameter tunning.
resamples <- train %>% vfold_cv(v = 10, repeats = 2)

# Define grid of parameters to test.
param_grid <- grid_regular(penalty(), mixture(),
                           levels = list(penalty = 100,
                                         mixture = 10))

model_4_wflow <- workflow() %>%
                 add_model(pen_log_mod) %>%
                 add_recipe(cleaning_recipe)
#tic()
#model_4_fit <- model_4_wflow %>%
               #tune_grid(resamples = resamples, grid = param_grid,
                         #metrics = metric_set(roc_auc))
#toc()
# model fit was saved an reloaded as a rds to save runtime.
```

```{r model 4 save and reload, include = F}
#save(file = "Figures/model_4_fit.rds", model_4_fit)
load("Figures/model_4_fit.rds")
```

```{r Exact Tunning Parameters}
model_4_best_params <- select_best(model_4_fit)
```

## Model Predictions and Evaluation
### Model 1
```{r Model 1 Train, cache = T}
model_1_train <- predict(log_Reg_fit, new_data = train, type = "prob")[2]
model_1_train$obs <- train$is_canceled
model_1_cut <- model_1_train %>%
  cutpointr(x = .pred_1, class = obs,
            method = maximize_metric, metric = cohens_kappa,
            pos_class = 1, direction = ">=")
model_1_train %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))


model_1_train_cmat <- model_1_train %>%
    conf_mat(truth = obs, estimate = pred.class)

model_1_train_metrics <- summary(model_1_train_cmat)
```

```{r Model 1 Test, cache = T}
model_1_test <- predict(log_Reg_fit, new_data = test, type = "prob")[2]
model_1_test$obs <- test$is_canceled

model_1_test %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))


model_1_test_cmat <- model_1_test %>%
    conf_mat(truth = obs, estimate = pred.class)

summary(model_1_test_cmat)
```

```{r Model 1 by Hotel, cache = T}
model_1_train %<>% mutate(hotel = train$hotel)
model_1_test %<>% mutate(hotel = test$hotel)

model_1_train_city <- model_1_train %>% filter(hotel == "City Hotel")
model_1_test_city <- model_1_test %>% filter(hotel == "City Hotel")
model_1_train_resort <- model_1_train %>% filter(hotel == "Resort Hotel")
model_1_test_resort <- model_1_test %>% filter(hotel == "Resort Hotel")


model_1_train_cmat_city <- model_1_train_city %>%
    conf_mat(truth = obs, estimate = pred.class)
model_1_train_metrics_city <- summary(model_1_train_cmat_city)

model_1_test_cmat_city <- model_1_test_city %>%
    conf_mat(truth = obs, estimate = pred.class)
model_1_test_metrics_city <- summary(model_1_test_cmat_city)

model_1_train_cmat_resort <- model_1_train_resort %>%
    conf_mat(truth = obs, estimate = pred.class)
model_1_train_metrics_resort <- summary(model_1_train_cmat_resort)

model_1_test_cmat_resort <- model_1_test_resort %>%
    conf_mat(truth = obs, estimate = pred.class)
model_1_test_metrics_resort <- summary(model_1_test_cmat_resort)
```

### Model 3
```{r Model 3 Train, cache = T}
model_3_train_city <- predict(model_3_fit_city,
    new_data = train_city, type = "prob")[2]

model_3_train_city$obs <- train_city$is_canceled

model_3_cut <- model_3_train_city %>%
  cutpointr(x = .pred_1, class = obs,
            method = maximize_metric, metric = cohens_kappa,
            pos_class = 1, direction = ">=")

model_3_train_city %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))

model_3_train_city_cmat <- model_3_train_city %>%
    conf_mat(truth = obs, estimate = pred.class)

model_3_train_city_metrics <- summary(model_3_train_city_cmat)

model_3_train_resort <- predict(model_3_fit_resort,
    new_data = train_resort, type = "prob")[2]

model_3_train_resort$obs <- train_resort$is_canceled

model_3_cut <- model_3_train_resort %>%
  cutpointr(x = .pred_1, class = obs,
            method = maximize_metric, metric = cohens_kappa,
            pos_class = 1, direction = ">=")
model_3_train_resort %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))

model_3_train_resort_cmat <- model_3_train_resort %>%
    conf_mat(truth = obs, estimate = pred.class)

model_3_train_resort_metrics <- summary(model_3_train_resort_cmat)
```

```{r Model 3 Test, cache = T}
model_3_test_city <- predict(model_3_fit_city,
                              new_data = test_city, type = "prob")[2]
model_3_test_city$obs <- test_city$is_canceled

model_3_test_city %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))

model_3_test_city_cmat <- model_3_test_city %>%
    conf_mat(truth = obs, estimate = pred.class)

model_3_test_city_metrics <- summary(model_3_test_city_cmat)

model_3_test_resort <- predict(model_3_fit_resort,
                              new_data = test_resort, type = "prob")[2]
model_3_test_resort$obs <- test_resort$is_canceled

model_3_test_resort %<>%
  mutate(pred.class = ifelse(.pred_1 >= model_1_cut$optimal_cutpoint, 1, 0)) %>%
  mutate(obs = as.factor(obs), pred.class = as.factor(pred.class))

model_3_test_resort_cmat <- model_3_test_resort %>%
    conf_mat(truth = obs, estimate = pred.class)

model_3_test_resort_metrics <- summary(model_3_test_resort_cmat)
```

## More Figures
```{r table 5}
table_5 <- cbind(model_1_train_metrics_city$.estimate,
                 model_1_test_metrics_city$.estimate,
                 model_1_train_metrics_resort$.estimate,
                 model_1_test_metrics_resort$.estimate,
                 model_3_train_city_metrics$.estimate,
                 model_3_test_city_metrics$.estimate,
                 model_3_train_resort_metrics$.estimate,
                 model_3_test_resort_metrics$.estimate)

table_5 <- as.data.frame(table_5) %>% slice(1:6)

rownames(table_5) <- c("Accuracy", "Kappa", "Sensitivity", "Specificity",
                       "PPV", "NPV")

colnames(table_5) <- c("Train", "Test",
                       "Train", "Test",
                       "Train", "Test",
                       "Train", "Test")


caption <- "Summary of model metrics"

table_5_latex <- table_5 %>%
  kbl(format = "latex", booktabs = T, longtable = T, caption = caption,
    linesep = "", escape = T, digits = 4) %>%
    kable_styling(full_width = F) %>%
    add_header_above(c(" ", "City" = 2, "Resort" = 2,
                            "City" = 2, "Resort" = 2)) %>%
    add_header_above(c(" ", "Combined Model" = 4, "Split Model" = 4))

table_5_latex

save(file = "Figures/table_5.rds", table_5_latex)
```

```{r Variable Importance Plot}
# Variable Importance
plot_4 <- log_Reg_glm %>%
          vip(num_features = 10) %+%
          aes(x = reorder(Variable, Importance), y = Importance,
              fill = Importance) +
          scale_fill_gradient() +
          labs(x = "") +
          theme_bw() + theme(legend.position = "none")

plot_4

save(file = "Figures/plot_4.rds", plot_4)
```

```{r Model 1 Coefficients}
caption <- "Final Model Coefficients"
model_1_coef <- log_Reg_glm %>% tidy()
# Fixes a latex table issue
model_1_coef$term[1] <- "Intercept"
model_1_coef %<>%
    kbl(format = "latex", booktabs = T, longtable = T, caption = caption,
      linesep = "", escape = T, digits = 3) %>%
      kable_styling(full_width = F)

model_1_coef

save(file = "Figures/model_1_coef.rds", model_1_coef)
```

```{r Roc Plots, eval = F}
model_1_roc <- layer_data(plot_roc(model_1_cut), 2) %>%
               select(x, y) %>% mutate(Model = "Model 1 Full")

model_1_roc <- layer_data(plot_roc(model_1_cut), 2) %>%
               select(x, y) %>% mutate(Model = "Model 1 Full")

```

```{r End Cluster, include = F}
stopCluster(cl)
```